Notes on Control Plane Inference

### 2021-03-31 inference.tex
add symbolic table encoding with havoc

% record fields
\newcommand{\matches}{\mathsf{matches}}
\newcommand{\action}{\mathsf{action}}
\newcommand{\actions}{\mathsf{actions}}
\newcommand{\keys}{\mathsf{keys}}
\newcommand{\data}{\mathsf{data}}
\newcommand{\id}{\mathsf{id}}

% commands
\newcommand{\assert}{\mathop{\mathsf{assert}}}
\newcommand{\assume}{\mathop{\mathsf{assume}}}
\newcommand{\apply}{\mathsf{apply}}
\newcommand{\choiceop}{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}
\newcommand{\choice}{\mathbin{\choiceop}}
\DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
\renewcommand{\choose}[2]{\mathop{\mathsf{havoc}~#1~\mathsf{from}~#2~\mathsf{in}}}
\newcommand{\SKIP}{\mathsf{skip}}

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
\newcommand{\nondet}{\mathsf{nondet}}
\newcommand{\ghostly}{\mathsf{ghostly}}

@@

\begin{figure}[htp]
  \[\begin{array}{lclll}
    \multicolumn{3}{l}{\Value} \\
    v & ::= & [n]_w  & \textit{Bitvector Literal} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\SymbRow} \\
    \rho & ::= & \left\{
    \begin{array}{l}
      \matches = \List[x];\\
      \action = x;\\
      \data = \List[x]
    \end{array}\right\} & \textit{Symbolic Row} & (\iota)\\
    \multicolumn{3}{l}{\BVExpr} \\
    e & ::= & v & \textit{Value} & (a,g,p,\iota)\\
      & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
      & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
      & \mid & \rho.\action & \textit{Symbolic Action} & (\iota) \\
      & \mid & \rho.\data[i] & \textit{Symbolic Action Data} & (\iota) \\
      & \mid & x & \textit{Action Data Variable} & (a^*) \\
      & \mid & e \binop e & \textit{Arbitrary Binary Operation} & (a,g,p,\iota) \\
    & \mid & \unop e & \textit{Arbitrary Unary Operation} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\Table} \\
    t & ::= & \left\{\begin{array}{l}
    t.\id = v; \\
    t.\keys = \List[h.f]; \\
    t.\actions = \List[\lambda \vec x.\,a]\\
    \end{array}\right\} & \textit{Table} & (p)\\
    \multicolumn{3}{l}{\BExpr} \\
    b & ::= & \FALSE  & \textit{Absurdity}& (a,g,p,\iota)\\
      & \mid & b \Rightarrow b & \textit{Implication} & (a,g,p,\iota)\\
      & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
    & \mid & \forall \rho \in t.~b & \textit{Row Universality} & (\iota)\\
    \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
    c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
      & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
      & \mid & \assume b & \textit{Assumption} & (g,p,\iota) \\
      & \mid & t.\apply() & \textit{Table Application} & (p)\\
      & \mid & c;c & \textit{Sequential Composition} & (a,g,p,\iota) \\
    & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
    & \mid & \choose \rho t c & \textit{Selection} & (\iota) \\
    \multicolumn{3}{l}{
      \begin{array}{ll}
        x \in \Var & h \in \Hdr\\
        n,w,i \in \mathbb{N} & f \in \Field
    \end{array}} & \textit{Sets }\\
  \end{array}
\]
\caption{Grammar of Commands $c \in \Cmd$. Legal action indicated by $a \in
  \Action$, legal instrumented command variants indicated by $\iota \in \Instr$,
  dataplane programs indicated by $p \in \Prog$, and table-free commands
  indicated by $g \in \GCL$.}
\label{fig:grammar}
\end{figure}

@@

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[?] : \Prog \to \Instr} \\
  (h.f := e)[?] && h.f := e \\
  (\assume b)[?] && \assume b \\
  (\assert b)[?] && \assert b\\
  (c_1;c_2)[?] && c_1[?];c_2[?]\\
  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
  (\choose \rho t c)[?] &&
  \textrm{\textcolor{red}{undefined}} \\
  (t.\apply())[?] &&
  \begin{array}[t]{l} \displaystyle
    \{ \\
    \quad \choose \rho t \;\{ \\
    \qquad \assume \rho.\matches = t.\keys; \\
    \qquad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left\{\begin{array}{l}
    \assume \rho.\action = i;\\
    t.\action[i](\rho.\data)
    \end{array}\right\}\\
    \quad \} \\
    \} \choice \{ \\
    \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
    \quad \SKIP \\
    \quad \}\\
    \}
  \end{array}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation}
  \label{fig:table-instrument}
\end{figure}

>>>>

% record fields
\newcommand{\matches}{\mathsf{matches}}
\newcommand{\action}{\mathsf{action}}
\newcommand{\actions}{\mathsf{actions}}
\newcommand{\keys}{\mathsf{keys}}
\newcommand{\data}{\mathsf{data}}
\newcommand{\params}{\mathsf{params}}
\newcommand{\id}{\mathsf{id}}

% commands
\newcommand{\assert}{\mathop{\mathsf{assert}}}
\newcommand{\assume}{\mathop{\mathsf{assume}}}
\newcommand{\apply}{\mathsf{apply}}
\newcommand{\choiceop}{\rotatebox[origin=c]{90}{$\sqsubset\!\!\!\sqsupset$}}
\newcommand{\choice}{\mathbin{\choiceop}}
\DeclareMathOperator*{\bigchoice}{\scalerel*{\choiceop}{\sum}}
\newcommand{\havoc}{\texttt{<?>}}
\newcommand{\SKIP}{\mathsf{skip}}

% math
\newcommand{\argmin}{\mathop{\textrm{argmin}}}
\newcommand{\WP}{\textit{wp}}
\newcommand{\satisfy}[3]{({#1,#3}) \models_{#2}}
\newcommand{\notsatisfy}[3]{({#1,#3})\not\models_{#2}}
\newcommand{\nondet}{\mathsf{nondet}}
\newcommand{\ghostly}{\mathsf{ghostly}}
\newcommand{\state}[1]{\llparenthesis#1\rrparenthesis}

@@

\begin{figure}[htp]
  \[\begin{array}{lclll}
    \multicolumn{3}{l}{\Value} \\
    v & ::= & [n]_w  & \textit{Bitvector Literal} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\SymbRow} \\
    \rho & ::= & \left\{
    \begin{array}{l}
      \matches = \List[x];\\
      \action = x;\\
      \data = \List[x]
    \end{array}\right\} & \textit{Symbolic Row} & (\iota)\\
    \multicolumn{3}{l}{\BVExpr} \\
    e & ::= & v & \textit{Value} & (a,g,p,\iota)\\
      & \mid & h.f & \textit{Header Access} & (a,g,p,\iota)\\
      & \mid & \havoc & \textit{Havoc} & (a,g,p,\iota) \\
      & \mid & \rho.\matches[i] & \textit{Symbolic Match} & (\iota)\\
      & \mid & \rho.\action & \textit{Symbolic Action} & (\iota) \\
      & \mid & \rho.\data[i] & \textit{Symbolic Action Data} & (\iota) \\
      & \mid & x & \textit{Action Data Variable} & (a^*) \\
      & \mid & e \binop e & \textit{Arbitrary Binary Operation} & (a,g,p,\iota) \\
    & \mid & \unop e & \textit{Arbitrary Unary Operation} & (a,g,p,\iota) \\
    \multicolumn{3}{l}{\Table} \\
    t & ::= & \left\{\begin{array}{l}
    t.\id = v; \\
    t.\keys = \List[h.f]; \\
    t.\actions = \List[\lambda \vec x.\,a]\\
    \end{array}\right\} & \textit{Table} & (p)\\
    \multicolumn{3}{l}{\BExpr} \\
    b & ::= & \FALSE  & \textit{Absurdity}& (a,g,p,\iota)\\
      & \mid & b \Rightarrow b & \textit{Implication} & (a,g,p,\iota)\\
      & \mid & e = e & \textit{Bitvector Equality} & (a,g,p,\iota)\\
    & \mid & \forall \rho \in t.~b & \textit{Row Universality} & (\iota)\\
    \multicolumn{4}{l}{c \in \Cmd, a \in \Action, g \in \GCL, p \in \Prog, \iota \in \Instr } \\
    c & ::=  & h.f := e & \textit{Assignment} & (g,p,\iota) \\
      & \mid & \assert b & \textit{Assertion} & (g,p,\iota)\\
      & \mid & \assume b & \textit{Assumption} & (g,p,\iota) \\
      & \mid & t.\apply() & \textit{Table Application} & (p)\\
      & \mid & c;c & \textit{Sequential Composition} & (a,g,p,\iota) \\
    & \mid & c \choice c & \textit{Nondeterministic Choice} & (g,p,\iota)\\
    \multicolumn{3}{l}{
      \begin{array}{ll}
        x \in \Var & h \in \Hdr\\
        n,w,i \in \mathbb{N} & f \in \Field
    \end{array}} & \textit{Sets }\\
  \end{array}
\]
\caption{Grammar of Commands $c \in \Cmd$. Legal action indicated by $a \in
  \Action$, legal instrumented command variants indicated by $\iota \in \Instr$,
  dataplane programs indicated by $p \in \Prog$, and table-free commands
  indicated by $g \in \GCL$.}
\label{fig:grammar}
\end{figure}

@@

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)[?] : \Cmd \to \Cmd} \\
  \multicolumn{3}{l}{(\cdot)[?] : \Prog \to \Instr} \\
  (h.f := e)[?] && h.f := e \\
  (\assume b)[?] && \assume b \\
  (\assert b)[?] && \assert b\\
  (c_1;c_2)[?] && c_1[?];c_2[?]\\
  (c_1 \choice c_2)[?] && c_1[?] \choice c_2[?]\\
  (t.\apply())[?] &&
  \begin{array}[t]{l} \displaystyle
    \{ \\
    \quad \rho.\matches := \havoc; \\
    \quad \rho.\action := \havoc;\\
    \quad \rho.\data := \havoc; \\
    \quad \assume |\rho.\matches| = |t.\keys|; \\
    \quad \assume \rho.\action < |t.\actions|; \\
    \quad \assume |\rho.\data| < |t.\action[\rho.\action].\params| \\
    \quad \assume \rho.\matches = t.\keys; \\
    \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left\{\begin{array}{l}
    \assume \rho.\action = i;\\
    t.\action[i](\rho.\data)
    \end{array}\right\}\\
    \} \choice \{ \\
    \quad \assume \forall \rho \in t. \rho.\matches \neq t.\keys; \\
    \quad \SKIP \\
    \}
  \end{array}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation}
  \label{fig:table-instrument}
\end{figure}

\begin{figure}[htp]
  \[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{\state\sigma_c : (\Value \to \Row^*) \to \BExpr} \\
  \state\sigma_c = \bigwedge_{ t \in \tables(c) } \bigvee_{r_i \in \sigma(t.\id)}
  \rho = r_i
  \end{array}
  \]
  \caption{Symbolic State Encoding}
  \label{fig:table-instrument}
\end{figure}
